First Season MWA EoR Power Spectrum Results at Redshift 7

### 2016-08-05 referee_response.txt

We have added an error plot in Section 3.3 (now Figure 3), along with a paragraph at the end of the section describing the structure.

======================================================================

---11th paragraph of Section 3.1. "At this stage we also average in frequency by a factor of two..." Is there a specific motivation for this? Or is it just a convenient way to reduce the data volume?

Yes, this is done to reduce data volume. We have added a sentence explaining this, and pointing out that it does not affect our sensitivity. (It does however, effect the kpar modes we access, which is the cause of the different k modes shown for FHD+eppsilon and RTS+CHIPS that you point out below.)

@@

Our understanding is that unpublished references should include the initials, while published references do not (http://journals.aas.org/authors/references.html#Unpublished). However, when double checking we noticed a couple unpublished references that were missing the initials, and have now added them.

======================================================================

---Penultimate paragraph of Section 4.4. What does "total residual power" mean? Can we be more precise? Is it the power integrated over the whole kperp-kpara plane? If so, the relevant k scales of integration need to be specified.

First, the 90% number turned out to be not quite right. When we returned to our calculation to answer this question, we realized we had mistakenly included the number from another test we did after “locking in” the analysis for this paper, where we included a Stokes Q model. The correct number should have been 70%, and has been updated in the paper (in all places it is mentioned).

To answer the question at hand, the total residual power is calculated by squaring and summing the residual image cube. This is a naturally weighted image, so the weighting is exactly the inverse of our error figure (now Fig. 2). While we include all scales measured by the instrument, it is dominated by the large perpendicular scales. This is now described in the paper.

### 2016-08-05 ms.tex
Finished addressing referee comments

\textbf{Figure~\ref{fig:example_2derror} shows the error level corresponding to the two 
dimensional power spectrum in Figure~\ref{fig:example_ps}. These errors
are found by propagating noise measured from calibrated visibilities through the entire
pipeline. The dense core of the MWA results in relatively low noise at low $k_\perp$.
As the $(u,v)$ density of the array decreases at longer baselines, the noise increases.
The noise is mostly constant across $k_{||}$ because the Fourier transform distributes
the noise in frequency to all modes equally \citep{Morales:2004}.
The exception is the second $k_{||}$ bin, which has slightly lower noise than the rest of the 
$k_{||}$ modes. 
As explained in Hazelton et al. (in prep.) this is a result of the Lomb-Scargle periodogram 
interacting with the Blackman-Harris window function we apply to the data. Most of the bins 
have the noise equally distributed between cosine and sine terms, with the exception of the 
$k_{||}=0$ bin where the cosine term carries all the information. Due to the correlations 
introduced by the Blackman-Harris window function the first non-zero $k_{||}$ mode is highly 
covariant with the bottom bin ? therefore the noise on this mode's cosine term is lower than 
other higher order cosines.
}

@@

For this iteration of analysis we only use the diffuse model for subtraction, and omit it for 
calibration. At the time of writing short baselines were not producing reliable results in the 
calibration loop. Instead we chose to omit the diffuse model and mask baselines shorter 
than 50 wavelengths for calibration purposes. However, we add the diffuse model and 
unmask short baselines when performing foreground subtraction, and have seen a 90\% 
decrease in total residual power -- a strong indication that our diffuse model improves our 
subtraction. Future iterations will incorporate the model in calibration and gradually improve 
the model itself.

@@

\begin{figure}
\begin{center}
\subfigure{\includegraphics[width=\columnwidth]{diffuse_diff_model.pdf}}
~
\subfigure{\includegraphics[width=\columnwidth]{diffuse_diff_residual.pdf}}
\caption{
Single polarization (E-W) power spectrum differences from the golden data set showing the 
effect of subtracting a diffuse foreground model. The calibration is identical with and without 
the diffuse model, so the dirty power spectra are identical. 
\textbf{In the top panel the model power spectrum difference is almost entirely negative 
(red), indicating that the diffuse model added a large amount of power to the foreground 
model. The bottom panel shows the residual power spectrum difference is positive in the 
wedge (blue), demonstrating a successful subtraction.}
While this first attempt at a diffuse model was rudimentary, it 
\textbf{successfully removed 90\% of the residual power. }
\label{fig:diffuse_diff}
}
\end{center}
\end{figure}

@@

\textbf{Diffuse emission model.} The diffuse model here contained no spectral information
and represented only the Stokes I polarization. Nevertheless subtracting this model removed
90\% of the residual power in the foreground wedge. Adding a spectral index and multiple
polarization components to the model will improve this subtraction even further.
As \textbf{\citet{Lenc:2016}} demonstrated, substantial polarized diffuse structure
exists in the EoR0 field. The contributions to Stokes Q and U depend on the ionospheric
conditions, which will be accounted for in future analyses to appropriately model this
emission. 
In addition the diffuse model will be extended to an all-sky model. It is evident from the
jackknives shown in Figure~\ref{fig:jackknife} that strong emission near the horizon is
leaking into our power spectra during off-zenith observations. By modeling and subtracting
this emission we may be able to recover these observations in future analyses.

@@

\textbf{Between 2013 and the time of writing, the MWA has collected over two thousand 
hours of data targeting the EoR fields. These observations include the three designated sky 
fields and two frequency bands. In addition, observations are underway to further 
characterize the sources in the instrument sidelobes. The second phase of the MWA
will be commissioned in the latter half of 2016, and will include two sets of highly redundant
cores of antennas. This hybrid configuration will enable a unique opportunity to study
the calibration techniques used in imaging and delay spectrum analyses, furthering
our understanding of the systematics limiting our current results.}

>>>>

\textbf{Figure~\ref{fig:example_2derror} shows the error level corresponding to the two 
dimensional power spectrum in Figure~\ref{fig:example_ps}. These errors
are found by propagating noise measured from calibrated visibilities through the entire
pipeline. The dense core of the MWA results in relatively low noise at low $k_\perp$.
As the $(u,v)$ density of the array decreases at longer baselines, the noise increases.
The noise is mostly constant across $k_{||}$ because the Fourier transform distributes
the noise in frequency to all modes equally \citep{Morales:2004}.
The exception is the second $k_{||}$ bin, which has slightly lower noise than the rest of the 
$k_{||}$ modes. 
As explained in Hazelton et al. (in prep.) this is a result of the Lomb-Scargle periodogram 
interacting with the Blackman-Harris window function we apply to the data. Most of the bins 
have the noise equally distributed between cosine and sine terms, with the exception of the 
$k_{||}=0$ bin where the cosine term carries all the information. Due to the correlations 
introduced by the Blackman-Harris window function the first non-zero $k_{||}$ mode is highly 
covariant with the bottom bin -- therefore the noise on this mode's cosine term is lower than 
other higher order cosines.
}

@@

For this iteration of analysis we only use the diffuse model for subtraction, and omit it for 
calibration. At the time of writing short baselines were not producing reliable results in the 
calibration loop. Instead we chose to omit the diffuse model and mask baselines shorter 
than 50 wavelengths for calibration purposes. However, we add the diffuse model and 
unmask short baselines when performing foreground subtraction.
\textbf{We compare the total residual power in the images by squaring and summing
the residual image cube from before and after the diffuse subtraction. 
This includes all scales measured by the instrument, but because the image cubes are
naturally weighted, the $\mathbf{k}$ space weighting is the inverse of 
Figure~\ref{fig:example_2derror}, meaning $k_\perp \lesssim 0.05$ h Mpc$^{-1}$ dominates the
average. We saw a 70\% reduction in residual power when including our diffuse model}
 -- a strong indication that our diffuse model improves our 
subtraction. Future iterations will incorporate the model in calibration and gradually improve 
the model itself.

@@

\begin{figure}
\begin{center}
\subfigure{\includegraphics[width=\columnwidth]{diffuse_diff_model.pdf}}
~
\subfigure{\includegraphics[width=\columnwidth]{diffuse_diff_residual.pdf}}
\caption{
Single polarization (E-W) power spectrum differences from the golden data set showing the 
effect of subtracting a diffuse foreground model. The calibration is identical with and without 
the diffuse model, so the dirty power spectra are identical. 
\textbf{In the top panel the model power spectrum difference is almost entirely negative 
(red), indicating that the diffuse model added a large amount of power to the foreground 
model. The bottom panel shows the residual power spectrum difference is positive in the 
wedge (blue), demonstrating a successful subtraction.}
While this first attempt at a diffuse model was rudimentary, it 
\textbf{successfully removed 70\% of the residual power. }
\label{fig:diffuse_diff}
}
\end{center}
\end{figure}

@@

\textbf{Diffuse emission model.} The diffuse model here contained no spectral information
and represented only the Stokes I polarization. Nevertheless subtracting this model removed
70\% of the residual power in the foreground wedge. Adding a spectral index and multiple
polarization components to the model will improve this subtraction even further.
As \textbf{\citet{Lenc:2016}} demonstrated, substantial polarized diffuse structure
exists in the EoR0 field. The contributions to Stokes Q and U depend on the ionospheric
conditions, which will be accounted for in future analyses to appropriately model this
emission. 
In addition the diffuse model will be extended to an all-sky model. It is evident from the
jackknives shown in Figure~\ref{fig:jackknife} that strong emission near the horizon is
leaking into our power spectra during off-zenith observations. By modeling and subtracting
this emission we may be able to recover these observations in future analyses.

@@

\textbf{Between 2013 and the time of writing, the MWA has collected over two thousand 
hours of data targeting the EoR fields. These observations include the three designated sky 
fields and two frequency bands. In addition, observations are underway to further 
characterize the sources in the instrument sidelobes. The second phase of the MWA
will be commissioned in the latter half of 2016, and will include two sets of highly redundant
cores of antennas. This hybrid configuration will enable a unique opportunity to study
the calibration techniques used in imaging and delay spectrum analyses, providing
another view of the systematics limiting our current results.}
